Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-27 paper/ijcai16.tex
paper: tolti tempi per r=5

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
%<<<<<<< HEAD
%varying the number of collected user responses. Here $u(\cdot)$ is the
%true unknown user utility, and $\vx^*$ is the solution recommended to
%the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
%
%The thick line is the median
%utility loss taken over all 20 users, while the shaded area represents the
%standard deviation.
%
%\stefano{plot average utility loss at the increase of dataset size}
%
%\stefano{plot average runtime at the increase of dataset size}
%
%\stefano{comment on usefulness of sparse weights}
%
%\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
%
%\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
%methods, which generate a single candidate option per iteration, {\sc
%SetMargin} constructs a whole {\em set} of options. We thus can not leverage
%standard query selection strategies \stefano{which ones?}. 
%
%The simplest solution is to query the user about all $k \choose 2$ pairs of
%configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
%number of queries does not grow large in practice.  An appealing alternative
%involves using a modified sorting algorithm, such as merge-sort, to order the
%set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
%approach is very susceptible to noise: a single inconsistent user response may
%affect all successive comparisons, leading to a cascading effect. While more
%robust strategies may be conceived, empirically we found that the simpler
%quadratic scheme tends to work reasonably well.
%
%\andrea{rewrite as footnote}
%=======
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and EUS) are
single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. \andrea{@stefano:
  possiamo riportare la mediana anche per i tempi?}  Our {\sc
  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
one iteration corresponds to a single pairwise query to the user.
Results for $r=3$ and $r=4$ are reported.  For dense weight vector
distributions (first two rows), our approach achieves results which
are indistinguishable from the competitors in a fraction of their
time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
  sicuro?}  while our algorithm can easily scale to much larger
datasets, as will be shown later on. For sparse weight vector
distributions (last two rows) our approach, in addition to being
substantially faster on each iteration, requires less queries in order
to reach optimal solutions. This is an expected result as the
sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
enforcing sparsity in the weight vectors, while none of the other
approaches are explicitly designed to favour sparse functions.

@@

This paper presented a maxmargin approach to generating a set of diverse examples.
Our approach, that can be seen as an extension of maxmargin learning to sets, is effective in the generation of a diverse set of items that can be used to ask informative queries to the user.
The main advantages of our elicitation method are 1) its aptitude for eliciting preferences and providing recommendations in large configuration problems  2) robustness with respect to erroneous preference input, and 3) its ability to support sparse utility functions.
In a number of experiments we showed how our method compare with state-of-the-art approaches.

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      and the Bayesian competitors on the $r=3$ (left) and $r=4$
      (right) datasets. Each row represents a different distribution
      for sampling the underlying user utility.  The x-axis reports
      the number of iterations (i.e. pairwise queries), while the
      y-axis reports the utility loss (first and third columns) and
      the cumulative time in seconds (second and fourth
      columns). Thick lines indicate median values over all users,
      while standard deviations are shown as shaded areas. Curves
      represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
      (green) and Viappiani EUS (gray) respectively. \andrea{@stefano remove median and average from y-axis}}
\end{figure}

>>>>

In Figure~\ref{fig:comparison} we report solution quality and timing
values for increasing number of collected user responses, for the
different competitors on each of the four different utility vector
distributions. Solution quality is measured in terms of utility loss:
%
$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
%
%<<<<<<< HEAD
%varying the number of collected user responses. Here $u(\cdot)$ is the
%true unknown user utility, and $\vx^*$ is the solution recommended to
%the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
%
%The thick line is the median
%utility loss taken over all 20 users, while the shaded area represents the
%standard deviation.
%
%\stefano{plot average utility loss at the increase of dataset size}
%
%\stefano{plot average runtime at the increase of dataset size}
%
%\stefano{comment on usefulness of sparse weights}
%
%\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
%
%\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
%methods, which generate a single candidate option per iteration, {\sc
%SetMargin} constructs a whole {\em set} of options. We thus can not leverage
%standard query selection strategies \stefano{which ones?}. 
%
%The simplest solution is to query the user about all $k \choose 2$ pairs of
%configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
%number of queries does not grow large in practice.  An appealing alternative
%involves using a modified sorting algorithm, such as merge-sort, to order the
%set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
%approach is very susceptible to noise: a single inconsistent user response may
%affect all successive comparisons, leading to a cascading effect. While more
%robust strategies may be conceived, empirically we found that the simpler
%quadratic scheme tends to work reasonably well.
%
%\andrea{rewrite as footnote}
%=======
where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
solution recommended to the user after the elicitation phase (see
Algorithm~\ref{alg:setmargin}). Computational cost is measured in
terms of cumulative time. Given that Guo and Viappiani (both QI and
EUS) are single-threaded, we disabled multi-threading when running our
algorithm in these comparisons. All experiments were run on a XXXXXX\andrea{@stefano}.
In all figures, thick lines indicate median values over the different
users, shaded areas indicate standard deviations. Our {\sc SetMargin}
algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
in green and Viappiani EUS in gray. For all algorithms, one iteration
corresponds to a single pairwise query to the user.  Results for $r=3$
and $r=4$ are reported.  For dense weight vector distributions (first
two rows), our approach achieves results which are indistinguishable
from the competitors in a fraction of their time.  Indeed, all
Bayesian approaches become quickly intractable for growing values of $r$,
%cannot scale over $r=4$ (for $r=5$, Guo did not
%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
while our algorithm can easily scale to much larger datasets, as will
be shown later on. For sparse weight vector distributions (last two
rows) our approach, in addition to being substantially faster on each
iteration, requires less queries in order to reach optimal
solutions. This is an expected result as the sparsification norm in
our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
vectors, while none of the other approaches are explicitly designed to
favour sparse functions.

@@

We presented a max-margin approach for generating a set of diverse
examples.  Our approach, that can be seen as an extension of maxmargin
learning to sets, is effective in the generation of a diverse set of
items that can be used to ask informative queries to the user.  The
main advantages of our elicitation method are 1) its aptitude for
eliciting preferences and providing recommendations in large
configuration problems 2) robustness with respect to erroneous
preference input, and 3) its ability to support sparse utility
functions.  In a number of experiments we showed how our method
compare with state-of-the-art approaches.

@@

\begin{figure}[b]
    \centering
    {\footnotesize
    \begin{tabular}{cccc}
        \hline
        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
        \\
        \hline \hline
        \multicolumn{4}{c}{{\sc Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Uniform}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
        \\
        \hline
        \multicolumn{4}{c}{{\sc Sparse Normal}}
        \\
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
        \\
        \hline
    \end{tabular}
    }
    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
      and the Bayesian competitors on the $r=3$ (left) and $r=4$
      (right) datasets. Each row represents a different distribution
      for sampling the underlying user utility.  The x-axis reports
      the number of iterations (i.e. pairwise queries), while the
      y-axis reports the utility loss (first and third columns) and
      the cumulative time in seconds (second and fourth
      columns). Thick lines indicate median values over all users,
      while standard deviations are shown as shaded areas. Curves
      represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
      (green) and Viappiani EUS (gray) respectively.}
\end{figure}
